---
title: "Wealth Inequality by Age Cohort"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(reldist)
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
library(ggplot2)
library(knitr)
library(kableExtra)
load(file = "../ct_demographics.Rda")
```

Wealth inequality, as measured by the Gini coefficient, decreases by age cohort in Meriden, in both 1850 and 1860. The graphs show the Gini coefficient for real property wealth in 1850 and 1860, and for total wealth in 1860, by age cohort.

The Gini coefficient is a measure of inequality, where 0 represents perfect equality and 1 represents perfect inequality.

```{r graphs, echo = FALSE}
ct_1860_hh_excl <- ct_1860_hh %>%
  dplyr::filter(town == "Meriden", family != 33643101)

age_cats <- c("20 - 29", "30 - 39", "40 - 49", "50 - 59", "60 and over")

# Define a function to calculate Gini coefficient by age category for a given set of household data
calc_gini_by_age <- function(data, town, wealth_col) {
  result <- c()
  
  for (age in age_cats) {
    gini_val <- data %>%
      dplyr::filter(town == town, AGE_CAT == age) %>%
      summarise(
        n = n(),
        gini = ifelse(n >= 5, reldist::gini({{ wealth_col }}), NA)
      ) %>%
      pull(gini)
    
    result <- c(result, gini_val)
  }
  
  return(tibble(gini = result))
}

# Calculate all four series
gini_1850_realprop <- calc_gini_by_age(ct_1850_hh, 'Meriden', FAMILY_REALPROP)
gini_1860_realprop <- calc_gini_by_age(ct_1860_hh, 'Meriden', FAMILY_REALPROP)
gini_1860_wealth <- calc_gini_by_age(ct_1860_hh, 'Meriden',FAMILY_WEALTH)
gini_1860_wealth_excl <- calc_gini_by_age(ct_1860_hh_excl, 'Meriden', FAMILY_WEALTH)

# Graph the inclusive results
gini_all <- tibble(
  age_cat = age_cats,
  realprop_1850 = gini_1850_realprop$gini,
  realprop_1860 = gini_1860_realprop$gini,
  wealth_1860 = gini_1860_wealth$gini
)

ggplot(gini_all, aes(x = age_cat)) +
  geom_line(aes(y = realprop_1850, color = "Real Property 1850"), group = 1) +
  geom_line(aes(y = realprop_1860, color = "Real Property 1860"), group = 1) +
  geom_line(aes(y = wealth_1860, color = "Total Wealth 1860"), group = 1) +
  ylim(0.65, 1) +
  labs(title = "Meriden Wealth Inequality",
       x = "Age Cohort",
       y = "Gini Coefficient") +
  scale_color_manual(values = c("blue", "red", "green")) +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "bottom" ) +  # Larger base
  theme(
    plot.title = element_text(hjust = 0.5, size = 24, face = "bold", margin = margin(b = 10)),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 12),
    plot.margin = margin(5.5, 5.5, 5.5, 5.5)
  )
```

The GINI coefficient for 1860 total wealth is correlated to total wealth as well as to age. Because wealth tends to increase with age, it's important to determine which factor better describes the variation in GINI coefficients

Analysis demonstrates that **age structure is a substantially better predictor of wealth inequality (GINI) than wealth itself**. While wealth and age are positively correlated, age alone explains the majority of variation in inequality, with wealth contributing only marginal unique explanatory power.

The following ANOVA shows how much variation in GINI each factor explains **sequentially** (controlling for previous variables):

```{r sequential-anova}
model <- lm(gini ~ pct_farm_1860 + age_1860 + wealth, data = factors)
anova_results <- anova(model)
total_ss <- sum(anova_results$`Sum Sq`)
anova_results$Prop_Var_Explained <- anova_results$`Sum Sq` / total_ss
anova_results$Cumulative_Prop <- cumsum(anova_results$Prop_Var_Explained)

# Format for presentation
anova_table <- anova_results %>%
  as.data.frame() %>%
  mutate(
    Variable = rownames(.),
    Prop_Var_Explained = scales::percent(Prop_Var_Explained, accuracy = 0.1),
    Cumulative_Prop = scales::percent(Cumulative_Prop, accuracy = 0.1),
    Significance = case_when(
      `Pr(>F)` < 0.001 ~ "***",
      `Pr(>F)` < 0.01 ~ "**",
      `Pr(>F)` < 0.05 ~ "*",
      TRUE ~ ""
    )
  ) %>%
  select(Variable, Df, `Sum Sq`, `Mean Sq`, `F value`, `Pr(>F)`, 
         Prop_Var_Explained, Cumulative_Prop, Significance)

kable(anova_table, 
      caption = "Table 1: Sequential ANOVA Results (Type I Sum of Squares)",
      digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

**Key finding:** Urbanization accounts for 41.0% of variation, age accounts for 9.2%, while wealth adds only 1.5% beyond these factors.

Wealth and age are positively correlated:

```{r age-wealth-correlation}
age_wealth_cor <- cor(factors$age_1860, factors$wealth)

cat("Pearson correlation (age × wealth): ", round(age_wealth_cor, 3))
```

Wealth tends to increase with age, suggesting potential confounding.

To isolate whether wealth explains inequality **beyond** what age already captures, we compare three models:

```{r model-comparison}
model_age_only <- lm(gini ~ age_1860, data = factors)
model_wealth_only <- lm(gini ~ wealth, data = factors)
model_both <- lm(gini ~ age_1860 + wealth, data = factors)

# Model summary table
model_summary <- data.frame(
  Model = c("Age only", "Wealth only", "Age + Wealth"),
  R_squared = c(
    summary(model_age_only)$r.squared,
    summary(model_wealth_only)$r.squared,
    summary(model_both)$r.squared
  ),
  Adj_R_squared = c(
    summary(model_age_only)$adj.r.squared,
    summary(model_wealth_only)$adj.r.squared,
    summary(model_both)$adj.r.squared
  ),
  AIC = c(AIC(model_age_only), AIC(model_wealth_only), AIC(model_both))
)

kable(model_summary %>%
        mutate(across(where(is.numeric), ~ round(., 4))),
      caption = "Table 2: Model Performance Comparison") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

**Observation:** Age alone (R² = `r round(summary(model_age_only)$r.squared, 3)`) explains substantially more variation than wealth alone (R² = `r round(summary(model_wealth_only)$r.squared, 3)`).

The critical test: Does including wealth **improve** a model that already includes age?

```{r anova-comparison}
anova_comparison <- anova(model_age_only, model_both)

# Create formatted table
anova_comp_table <- data.frame(
  Model = c("Age only", "Age + Wealth"),
  Res_DF = anova_comparison$Res.Df,
  RSS = anova_comparison$RSS,
  DF_Change = c(NA, anova_comparison$Df[2]),
  SS_Change = c(NA, anova_comparison$`Sum of Sq`[2]),
  F_Value = c(NA, anova_comparison$F[2]),
  P_Value = c(NA, anova_comparison$`Pr(>F)`[2])
)

kable(anova_comp_table %>%
        mutate(across(where(is.numeric), ~ round(., 4))),
      caption = "Table 3: ANOVA Comparison - Does Wealth Improve the Age Model?",
      na = "—") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# Calculate additional R²
rss_age_only <- anova_comparison$RSS[1]
rss_both <- anova_comparison$RSS[2]
total_ss <- sum(anova(model)$`Sum Sq`)
r2_age <- 1 - (rss_age_only / total_ss)
r2_both <- 1 - (rss_both / total_ss)
r2_added <- r2_both - r2_age

cat("\n\n**Additional R² contributed by wealth:** ", 
    scales::percent(r2_added, accuracy = 0.1))
```

1. **Age structure is a strong predictor** of wealth inequality, explaining approximately 42.5% of variation (with urbanization)

2. **Wealth adds minimal unique explanatory power** - only 2.7% beyond what age already explains

3. **The age-wealth correlation** (r = `r round(age_wealth_cor, 3)`) suggests that wealth's apparent effect is largely mediated by age

4. **Substantive interpretation**: Demographic composition matters for inequality in ways that go beyond simple wealth accumulation. Policies targeting age structure or life-cycle effects may be more relevant to inequality than direct wealth redistribution.