---
title: "Correlation between votes for different offices"
output: html_document
---
  
```{r setup, include=FALSE}
library(dplyr)
library(magrittr)
wd <- getwd()
setwd("../")
source("./global.R", local = TRUE)

source("./results_utils.R", local = TRUE)
source("./census_utils.R", local = TRUE)
source("./present.R", local = TRUE)
governor_results <- read_results(results_file, alias_map, party_assignments, "Governor", 1849, 1857)
lt_governor_results <- read_results(results_file, alias_map, party_assignments, "Lieutenant Governor", 1849, 1857)
secretary_results <- read_results(results_file, alias_map, party_assignments, "Secretary of the State", 1849, 1857)
treasurer_results <- read_results(results_file, alias_map, party_assignments, "Treasurer", 1849, 1857)
probate_results <- read_probate_results(results_file, alias_map, party_assignments, 1849, 1857)
setwd(dirname(wd))
```

By Connecticut law, voters deposited ballots for probate judge in a separate ballot box from the one into which they deposited ballots for statewide offices. Almost all voters used printed ballots that had been handed to them by political operatives who worked the crowd in front of the voting window.

The close correlation of the votes for statewide offices demonstrates the tendency by voters to accept a ballot from a representative of one of the parties. The much weaker correlation of probate judge votes with those for statewide offices suggests either that operatives offered a combination of ballots that voters would be most likely to accept or that voters accepted ballots from different operatives for different offices. In either case, the weaker correlation for probate judge votes suggests that voters were less consistent in their party preferences than the "party period" model would suggest.

Because of the large number of candidates for probate judge, I have not attempted to identify them by party. Instead, I have assigned to the top three vote-getters in a given town and year the parties of the gubernatorial candidate in the same position for that town and year. This assumption produces the highest possible correlation between probate judge and gubernatorial votes, making the low correlations that are calculated all the more striking.

```{r table, echo = FALSE}
# Keep only those rows that have results for all five offices
governor_results <- semi_join(governor_results, probate_results, by = c("town", "yr")) %>%
  semi_join(lt_governor_results, by = c("town", "yr")) %>%
  semi_join(secretary_results, by = c("town", "yr")) %>%
  semi_join(treasurer_results, by = c("town", "yr"))
lt_governor_results <- semi_join(lt_governor_results, probate_results, by = c("town", "yr")) %>%
  semi_join(governor_results, by = c("town", "yr")) %>%
  semi_join(secretary_results, by = c("town", "yr")) %>%
  semi_join(treasurer_results, by = c("town", "yr"))
secretary_results <- semi_join(secretary_results, probate_results, by = c("town", "yr")) %>%
  semi_join(governor_results, by = c("town", "yr")) %>%
  semi_join(lt_governor_results, by = c("town", "yr")) %>%
  semi_join(treasurer_results, by = c("town", "yr"))
treasurer_results <- semi_join(treasurer_results, probate_results, by = c("town", "yr")) %>%
  semi_join(governor_results, by = c("town", "yr")) %>%
  semi_join(lt_governor_results, by = c("town", "yr")) %>%
  semi_join(secretary_results, by = c("town", "yr"))
probate_results <- semi_join(probate_results, governor_results, by = c("town", "yr")) %>%
  semi_join(lt_governor_results, by = c("town", "yr")) %>%
  semi_join(secretary_results, by = c("town", "yr")) %>%
  semi_join(treasurer_results, by = c("town", "yr"))

corr_df <- corr_matrix_by_party(governor_results, lt_governor_results, secretary_results, treasurer_results, probate_results, "Democrat")
knitr::kable(corr_df,
  format = "html"
) %>%
  kableExtra::add_header_above(c("Correlation of Democratic vote tallies for statewide offices, 1849-1857" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)

corr_df <- corr_matrix_by_party(governor_results, lt_governor_results, secretary_results, treasurer_results, probate_results, "Whig")
knitr::kable(corr_df,
  format = "html"
) %>%
  kableExtra::add_header_above(c("Correlation of Whig vote tallies for statewide offices, 1849-1857" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)

corr_df <- corr_matrix_by_party(governor_results, lt_governor_results, secretary_results, treasurer_results, probate_results, "Free_Soil")
knitr::kable(corr_df,
  format = "html"
) %>%
  kableExtra::add_header_above(c("Correlation of Free Soil vote tallies for statewide offices, 1849-1857" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)

corr_df <- corr_matrix_by_party(governor_results, lt_governor_results, secretary_results, treasurer_results, probate_results, "Temperance")
knitr::kable(corr_df,
  format = "html"
) %>%
  kableExtra::add_header_above(c("Correlation of Temperance vote tallies for statewide offices, 1849-1857" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)

corr_df <- corr_matrix_by_party(governor_results, lt_governor_results, secretary_results, treasurer_results, probate_results, "Know_Nothing")
knitr::kable(corr_df,
  format = "html"
) %>%
  kableExtra::add_header_above(c("Correlation of Know Nothing vote tallies for statewide offices, 1849-1857" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)

corr_df <- corr_matrix_by_party(governor_results, lt_governor_results, secretary_results, treasurer_results, probate_results, "Republican")
knitr::kable(corr_df,
  format = "html"
) %>%
  kableExtra::add_header_above(c("Correlation of Republican vote tallies for statewide offices, 1849-1857" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)
```
