---
title: "Connecticut Gubernatorial Party Transitions: 1851-1857"
output: html_document
---

```{r setup, include=FALSE}
library(sankeyD3plus)
wd <- getwd()
setwd("../")
source("./global.R", local = TRUE)
source("./betas.R", local = TRUE)
source("./present.R", local = TRUE)
source("./inference_utils.R", local = TRUE)
load("./sankey.Rda")
load("./results.Rda")
load("./1852_covariate_inference.Rda")
load("./1853_covariate_inference.Rda")
load("./1854_covariate_inference.Rda")
load("./1855_covariate_inference.Rda")
load("./1856_covariate_inference.Rda")
load("./1857_covariate_inference.Rda")
setwd(dirname(wd))
```

```{r sankey, echo=FALSE}
sankey_widget
```

Transitions less than 0.4% are not shown. Hover over the flows to see exact percentages and the degree of uncertainty, represented by the maximun R-hat value; hover over the nodes to see total percentages for each party in a given year and to highlight the flows into and out of that party from the previous year and into the following year.

Percentages may differ slightly from those in the transition tables below due to rounding.

```{r tables, echo = FALSE}
estimate <- choose_final_estimate(variability_results.52)
betas.52 <- estimate$estimates
kable.52 <- construct_contingency(
  results.52,
  betas.52,
  1851,
  1852
)
estimate <- choose_final_estimate(variability_results.53)
betas.53 <- estimate$estimates
kable.53 <- construct_contingency(
  results.53,
  betas.53,
  1852,
  1853
)
estimate <- choose_final_estimate(variability_results.54)
betas.54 <- estimate$estimates
kable.54 <- construct_contingency(
  results.54,
  betas.54,
  1853,
  1854
)
estimate <- choose_final_estimate(variability_results.55)
betas.55 <- estimate$estimates
kable.55 <- construct_contingency(
  results.55,
  betas.55,
  1854,
  1855
)
estimate <- choose_final_estimate(variability_results.56)
betas.56 <- estimate$estimates
kable.56 <- construct_contingency(
  results.56,
  betas.56,
  1855,
  1856
)
estimate <- choose_final_estimate(variability_results.57)
betas.57 <- estimate$estimates
kable.57 <- construct_contingency(
  results.57,
  betas.57,
  1856,
  1857
)
knitr::kable(kable.52,
  format = "html") %>%
  kableExtra::add_header_above(c("Estimated Voter Transitions from 1851 to 1852" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)
knitr::kable(kable.53,
  format = "html") %>%
  kableExtra::add_header_above(c("Estimated Voter Transitions from 1852 to 1853" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)
knitr::kable(kable.54,
  format = "html") %>%
  kableExtra::add_header_above(c("Estimated Voter Transitions from 1853 to 1854" = 7),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)
knitr::kable(kable.55,
  format = "html") %>%
  kableExtra::add_header_above(c("Estimated Voter Transitions from 1854 to 1855" = 6),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)
knitr::kable(kable.56,
  format = "html") %>%
  kableExtra::add_header_above(c("Estimated Voter Transitions from 1855 to 1856" = 7),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)
knitr::kable(kable.57,
  format = "html") %>%
  kableExtra::add_header_above(c("Estimated Voter Transitions from 1856 to 1857" = 5),
                               align = "left",
                               font_size = "larger") %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = FALSE)
```